---
title: "US Arrests Data Transformations"
author: "Katie, Sarah, Meerim"
date: "February 29, 2016"
output: pdf_document
---

source('~/Documents/Hertie 2016/Collaborative Social Science Data/Assignment1repo/USArrests_SourceFile.R')


```{r, echo=FALSE}
#use "USArrests" data set within R
data("USArrests")
?USArrests
```


## Violent Crime Rates by US State

This data set contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. Also given is the percent of the population living in urban areas.

```{r, echo=FALSE}
summary(USArrests)
```

The data examined and transformed is a data frame with 50 observations on 4 variables:

```{r, echo=FALSE}
names(USArrests)
```

## Frequency of Violent Crime Arrests by Crime Type

```{r, echo=FALSE}
old.par <- par(mfrow=c(1, 3))
hist(USArrests$Murder, main = 'Murder Arrests for the US states in 1973', xlab = 'Murder Arrests per 100,000 residents')
hist(USArrests$Rape, main = 'Rape Arrests for the US states in 1973', xlab = 'Rape Arrests per 100,000 residents')
hist(USArrests$Assault, main = 'Assault Arrests for the US states in 1973', xlab = 'Assault Arrests per 100,000 residents')
par(old.par)
```

##Overview of Violent Crime Arrests
```{r, echo=FALSE}
boxplot(USArrests)
```

## Violent Arrests and Urban Populations

Violent arrests plotted against the percent of the urban population

```{r USArrests, echo=FALSE}
old.par <- par(mfrow=c(1, 3))
#plot of Urban Population vs. Murders
plot(x = USArrests$UrbanPop, y = USArrests$Murder,xlab = "Urban Population (%)", ylab = "Murder (Arrests per 100,000)", pch=0)
title("Murder Arrest Rates")
fit <- lm(USArrests$Murder ~ USArrests$UrbanPop)   # Murder 'as a linear function of' Urban Population
abline(fit)

#plot of Urban Population vs. Rapes
plot(x = USArrests$UrbanPop, y = USArrests$Rape,xlab = "Urban Population (%)", ylab = "Rape (Arrests per 100,000)", pch=1)
title("Rape Arrest Rates")
fit <- lm(USArrests$Rape ~ USArrests$UrbanPop)   # Rape 'as a linear function of' Urban Population
abline(fit)

#plot of Urban Population vs. Assault
plot(x = USArrests$UrbanPop, y = USArrests$Assault,xlab = "Urban Population (%)", ylab = "Assault (Arrests per 100,000)", pch=2)
title("Assault Arrest Rates")
fit <- lm(USArrests$Assault ~ USArrests$UrbanPop)   # Assault 'as a linear function of' Urban Population
abline(fit)
par(old.par)
```


## Correlations between Urban Population and Violent Crime Rate

```{r, echo=FALSE}
cor.test(log(USArrests$UrbanPop), USArrests$Assault)
cor.test(log(USArrests$UrbanPop), USArrests$Rape)
cor.test(log(USArrests$UrbanPop), USArrests$Murder)
```




