---
title: "Motor Trend Cars"
author: "Katie, Sarah, Meerim"
date: "March 4, 2016"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source('~/GitHub/Assignment1repo/SourceFile.R')

```

## Data

This data is from Motor Trend magazine in 1974. It presents data for 32 automobiles regarding performance  and design.

Variables:

```{r summary mtcars, echo=FALSE}

names(mtcars)

```


Explanation of variable names: mpg is miles per gallon; 
cyl is number of cylinders; 
disp is displacement (in cubic inches); 
hp is horsepower; 
drat is rear axle ratio; 
wt is weight (in 1000 pounds); 
qsec is 1/4 mile time; 
vs is engine type (0 = vertical/1 = straight); 
am is transmission type (0 = automatic, 1 = manual); 
gear is number of forward gears; and 
carb is number of carburetors


``` {r labels, echo = FALSE}
# Labelling variables
var.labels = c(mpg="Miles per gallon", cyl="Number of cylinders", disp="Displacement", hp="Horsepower", drat="Axle ratio", wt="Weight", qsec="1/4 mile time", vs="V/S", am="Transmission", gear="Number of forward gears", carb="Number of carburetors")

label(mtcars) = lapply(names(var.labels), 
function(x) label(mtcars[,x]) = var.labels[x])

```


Basic statistics
```{r descr, echo=FALSE}
options(digits=2)
stat.desc(mtcars, basic = TRUE, desc=FALSE) 
```


There are no null or missing values. Displacement and horsepower have wide ranges. 


Means
``` {r means, echo=FALSE}
for (i in 1:length(names(mtcars))) {
  mtcars[, i] %>%
    mean(,na.rm=FALSE) %>%
    round(digits=2) %>%
    paste(names(mtcars)[i], ., '\n') %>%  
    cat()
}
```


The average miles per gallon is about 20, the average horsepower is about 147, and the average quarter-mile time is about 18 seconds. For engine and transmission types, 44% have vertical-style engines and 41% have automatic transmissions. Other characteristics of the data will be further explored in the plots. 


# Plots

### Boxplot distributions: MPG

```{r am boxplots, echo=FALSE}
# encode "am" as a vector (or category) and label 0/1 as automatic/manual
mtcars$am <- factor(mtcars$am,levels=c(0,1), labels=c("Automatic","Manual"))
# now plot mpg as a boxplot for the two transmission types on one graph, coloring each factor using "fill"
qplot(mtcars$am, mtcars$mpg, geom = "boxplot", main = "MPG by Transmission Type", xlab = "Transmission", ylab = "MPG", fill = factor(mtcars$am)) + guides(fill=FALSE) +scale_fill_manual(values=c("cyan1", "cyan3"))
# repeat encoding and plotting process for "vs""
mtcars$vs <- factor(mtcars$vs, levels =c(0,1), labels=c("Vertical", "Straight"))
qplot(mtcars$vs, mtcars$mpg, geom = "boxplot", main = "MPG by Engine Type", xlab = "Engine", ylab = "MPG", fill = factor(mtcars$vs)) + guides(fill=FALSE) +scale_fill_manual(values=c("cyan2", "cyan4"))
```

Manual transmissions and straight engines in this data set generally have better gas mileage. 


### Boxplot distributions: Horsepower, Weight and Displacement
```{r boxplots, echo=FALSE, message=FALSE}
# set up a graphic with multiple charts (1 row, 3 columns) using the mtcars data
attach(mtcars)
par(mfrow=c(1,3)) 
# create simple boxplots, with labels and colors
boxplot(hp, col="tomato3", xlab="Horsepower")
boxplot(wt, col="sienna1", xlab="Weight", ylab="in 1000 pounds")
boxplot(disp, col="tomato", xlab="Displacement", ylab="in cubic inches")
```

All three variables vary widely, but displacement has an especially wide distribution. 


### Histograms of gear, carburetor and cylinder Variables
``` {r histograms, echo=FALSE, message=FALSE}
# set up a graphic with multiple charts (1 row, 3 columns) 
attach(mtcars)
par(mfrow=c(1,3))
# create a vector telling r the bins to use for a histogram
gearbins <- c(1,2,3,4,5)
# create histograms, with labels and colors
hist(gear, xlab = "Gears", breaks=gearbins, col="powderblue", main=paste(var.labels["gear"]))
hist(carb, xlab = "Carburetors", col="lightskyblue", main=paste(var.labels["carb"]))
hist(cyl, xlab = "Cylinders", col="skyblue2", main=paste(var.labels["cyl"]))
```

Most cars had 3 or 4 gears, 1 or 3 carburetors, and 4 or 8 cylinders. 


### Scatterplots showing correlation of variables with MPG, with lines of best fit and Lowess lines
``` {r scatterplots, echo=FALSE}
# set up a graphic with multiple charts (2 rows, 2 columns) using the mtcars data
par(mfrow=c(2,2))
# create scatterplot of wt and mpg with labels
plot(wt,mpg, main="Weight and MPG", xlab = "Weight") 
# add line of best fit
abline(lm(mpg~wt), col="palegreen4")
# add lowess line
lines(lowess(wt,mpg, f=.9), col="palegreen")
# repeat for other variables
plot(hp, mpg, main="Horsepower and MPG", xlab="Horsepower")
abline(lm(mpg~hp), col="palegreen4")
lines(lowess(hp,mpg, f=.9), col="palegreen") 
plot(gear,mpg, main="Gears and MPG", xlab = "Gears") 
abline(lm(mpg~gear), col="palegreen4")
lines(lowess(gear,mpg, f=.9), col="palegreen")
plot(cyl,mpg, main="Cylinders and MPG", xlab = "Cylinders") 
abline(lm(mpg~cyl), col="palegreen4")
lines(lowess(cyl,mpg, f=.9), col="palegreen")
```

Weight, horsepower and cylinders are all negatively correlated with gas mileage. The number of gears does not appear to have a strong effect, though the best-fit line shows a slightly positve correlation. 


### Scatterplots showing correlation of variables with quarter-mile time, with lines of best fit and Lowess lines
``` {r more scatterplots, echo=FALSE}
# set up a graphic with multiple charts (2 rows, 2 columns) using the mtcars data
par(mfrow=c(2,2))
# create more scatterplots as above, with labels, lines of best fit and lowess lines
plot(wt,qsec, main="Weight and Quarter-mile", xlab = paste(var.labels["wt"]), ylab="in seconds") 
abline(lm(qsec~wt), col="mediumorchid4")
lines(lowess(wt,qsec, f=.9), col="mediumorchid")
plot(hp, qsec, main="Horsepower and Quarter-mile", xlab=paste(var.labels["hp"]), ylab="in seconds")
abline(lm(qsec~hp), col="mediumorchid4")
lines(lowess(hp,qsec, f=.9), col="mediumorchid")
plot(disp, qsec, main="Displacement and Quarter-mile", xlab=paste(var.labels["disp"]), ylab="in seconds")
abline(lm(qsec~disp), col="mediumorchid4")
lines(lowess(disp,qsec, f=.9), col="mediumorchid")
plot(mpg, qsec, main="MPG and Quarter-mile", xlab=paste(var.labels["mpg"]), ylab="in seconds")
abline(lm(qsec~mpg), col="mediumorchid4")
lines(lowess(mpg,qsec, f=.9), col="mediumorchid")
```

Horsepower and displacement are negatively correlated with the quarter-mile time, that is, cars with more horsepower or a larger engine have faster times. Weight shows a very slight negative correlation, too. Gas mileage and quarter-mile time are positively correlated: the better the gas mileage, the slower the time. 