---
title: "US Arrests Data Transformations"
author: "Katie, Sarah, Meerim"
date: "February 29, 2016"
output: pdf_document
---

source('~/Documents/Hertie 2016/Collaborative Social Science Data/Assignment1repo/USArrests_SourceFile.R')


```{r, echo=FALSE}
#use "USArrests" data set within R
data("USArrests")
?USArrests
```


## Violent Crime Rates by US State

This data set contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. Also given is the percent of the population living in urban areas.

```{r, echo=FALSE}
summary(USArrests)
```

The data examined and transformed is a data frame with 50 observations on 4 variables:

```{r, echo=FALSE}
names(USArrests)
```

## Frequency of Violent Crime Arrests by Crime Type

*Arrests for the US states in 1973 Per 100,000 Residents*
```{r, echo=FALSE}
old.par <- par(mfrow=c(1, 3))
hist(USArrests$Murder, main = 'Murder Arrests', xlab = 'Murder Arrests per 100,000 residents', col="powderblue")
hist(USArrests$Rape, main = 'Rape Arrests', xlab = 'Rape Arrests per 100,000 residents', col="lightskyblue")
hist(USArrests$Assault, main = 'Assault Arrests', xlab = 'Assault Arrests per 100,000 residents', col="skyblue2")
par(old.par)
```

*Former Southern States
```{r, echo=FALSE}
USArrests$state <- rownames(USArrests)
USArrests$dixie <- USArrests$state
USArrests$dixie <- 0
USArrests$dixie[which(USArrests$state=="Florida")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Georgia")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="North Carolina")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="South Carolina")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Mississippi")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Alabama")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Texas")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Virginia")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Arkansas")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Tennessee")] <- 1 #former Confederate State
USArrests$dixie[which(USArrests$state=="Louisiana")] <- 1 #former Confederate State
# Descriptive Statistics: Dixie Murder
par(mfrow=c(1,2))
hist(USArrests$Murder[USArrests$dixie==0],
     main="Rate North",
     col=(c("firebrick1")),
     xlab="Murder Arrests/100 000", ylab="Frequency")
hist(USArrests$Murder[USArrests$dixie==1],
     main="Rate South",
     col=(c("mediumspringgreen")),
     xlab="Murder Arrests/100 000", ylab="Frequency")
hist(USArrests$Murder[USArrests$dixie==0],
     main="Rate North",
     col=(c("firebrick1")),
     xlab="Assault Arrests/100 000", ylab="Frequency")
hist(USArrests$Assault[USArrests$dixie==1],
     main="Rate South",
     col=(c("mediumspringgreen")),
     xlab="Assault Arrests/100 000", ylab="Frequency")
hist(USArrests$Murder[USArrests$dixie==0],
     main="Rate North",
     col=(c("firebrick1")),
     xlab="Rape Arrests/100 000", ylab="Frequency")
hist(USArrests$Murder[USArrests$dixie==1],
     main="Rate South",
     col=(c("mediumspringgreen")),
     xlab="Rape Arrests/100 000", ylab="Frequency")
par(mfrow=c(1,1))
list(Mmean <- tapply(USArrests$Murder, USArrests$dixie, mean))
list(Amean <- tapply(USArrests$Assault, USArrests$dixie, mean))
list(Rmean <- tapply(USArrests$Rape, USArrests$dixie, mean))
```

##Overview of Violent Crime Arrests

All three variables vary widely, but **assault** has an especially wide distribution. There are far more assaults arrests per 100,000 than the other types of violent crime.

*Distribution of Arrests for the US States in 1793*

```{r, echo=FALSE}
attach(USArrests)
par(mfrow=c(1,3)) 
# create simple boxplots, with labels and colors
boxplot(USArrests$Murder, col="tomato3", xlab="Murder", ylab="Arrests per 100,000 Reisdents")
boxplot(USArrests$Assault, col="sienna1", xlab="Assault")
boxplot(USArrests$Rape, col="tomato", xlab="Rape")
```



## Violent Arrests and Urban Populations

Violent arrests plotted against the percent of the urban population. Murder, assault and rape are all positively correlated with the percentage of urban population. The murder does not appear to have a strong effect, though the best-fit line shows a slightly positve correlation.

```{r USArrests, echo=FALSE}
old.par <- par(mfrow=c(1, 3))
#plot of Urban Population vs. Murders
plot(x = USArrests$UrbanPop, y = USArrests$Murder,xlab = "Urban Population (%)", ylab = "Murder (Arrests per 100,000)", pch=0)
title("Murder Arrest Rates")
fit <- lm(USArrests$Murder ~ USArrests$UrbanPop)   # Murder 'as a linear function of' Urban Population
abline(fit,  col="tomato")

#plot of Urban Population vs. Rapes
plot(x = USArrests$UrbanPop, y = USArrests$Rape,xlab = "Urban Population (%)", ylab = "Rape (Arrests per 100,000)", pch=1)
title("Rape Arrest Rates")
fit <- lm(USArrests$Rape ~ USArrests$UrbanPop)   # Rape 'as a linear function of' Urban Population
abline(fit, col="tomato")

#plot of Urban Population vs. Assault
plot(x = USArrests$UrbanPop, y = USArrests$Assault,xlab = "Urban Population (%)", ylab = "Assault (Arrests per 100,000)", pch=2)
title("Assault Arrest Rates")
fit <- lm(USArrests$Assault ~ USArrests$UrbanPop)   # Assault 'as a linear function of' Urban Population
abline(fit, col="tomato")
par(old.par)
```


## Correlations between Urban Population and Violent Crime Rate


The correlation between the percentage of urban population and assault is **0.26**, indicating a positive correlation.

```{r, echo=FALSE}
cor.test(log(USArrests$UrbanPop), USArrests$Assault)
```

The correlation between the percentage of urban population and rape is **0.41**, indicating a strong, positive correlation.

```{r, echo=FALSE}
cor.test(log(USArrests$UrbanPop), USArrests$Rape)
```

The correlation between the percentage of urban population and murder is **0.08**, indicating a weakly positive correlation.

```{r, echo=FALSE}
cor.test(log(USArrests$UrbanPop), USArrests$Murder)
```




